All Your Graphs Belong to Us

### 2012-07-20 doc/asplos13/method.tex
refinements to UTS methodology as in-memory. Still need to discuss representation of irregular parallelism, as well as tree search

One question that must be answered when making a comparison between the Cray XMT and a typical HPC cluster is what is a fair comparison -- these systems are quite different.  Three options immediately come to mind: equal number of processing cores, equal number of network interfaces, and equal dollars.  We discounted the last option fairly quickly, because it isn't a lasting data-point and, as will be evident later, would be hugely unfair to the XMT.  The first option, cores, has some merit, but Grappa is designed for applications that have no locality in their computation.  This means almost all of their memory accesses are remote.  Hence, the factor that limits their performance is not processing, but communicating.  Hence, we have chosen the middle option, network interfaces as the way to normalize across the XMT and our cluster.  Each processor in the XMT system has its own network interface to access shared memory.  In the HPC cluster we use, each processor (which contains 12 cores), has a single Infiniband interface.  Hence, for our results we scale up XMT processors one for one with full system nodes.

@@

\subsection{Applications}
%Traversal of an unbalanced, unpartitioned tree captures the foundations for implementing most forms of irregular parallelism. 
%Recursion is found in every irregular divide and conquer algorithm. Loops, even regular ones, must be dynamically scheduled to get good load balance when run on a large, non-uniform system. Nested parallelism involves fine-grain creation of dynamic amounts of work. Dataflow..

\paragraph{Unbalanced tree search in-memory (UTS-Mem)}. Unbalanced
Search Tree (UTS) is a benchmark
for evaluating the programmability and performance of systems for
parallel applications that require dynamic load balancing
\cite{Olivier:uts2006}. It involves traversing an unbalanced implicit
tree: at each vertex, its number of children is sampled from some
probability distribution, and this number of new nodes are added to a
work queue to be visited. While this benchmark captures irregular,
dynamic parallelism, we want to evaluate performance on algorithms
that touch existing data.

We augment UTS by using the traversal to
create a large tree in memory, and then we traverse the in-memory
tree. We call this benchmark UTS-Mem, and the timed portion is this traversal of the in-memory tree. The assumption is that in general the size of each subtree is unknown until that subtree has actually been traversed.

>>>>

One question that must be answered when making a comparison between
the Cray XMT and a typical HPC cluster is what is a fair comparison --
these systems are quite different.  Three options immediately come to
mind: equal number of processing cores, equal number of network
interfaces, and equal dollars.  \checkme{We discounted the last option fairly
quickly, because it isn't a lasting data-point and, as will be evident
later, would be hugely unfair to the XMT.}  The first option, cores,
has some merit, but Grappa is designed for applications that have no
locality in their computation.  This means almost all of their memory
accesses are remote.  Hence, the factor that limits their performance
is not processing, but communicating.  Hence, we have chosen the
middle option, network interfaces as the way to normalize across the
XMT and our cluster.  Each processor in the XMT system has its own
network interface to access shared memory.  In the HPC cluster we use,
each processor (which contains up to 32 cores), has a single Infiniband interface.  Hence, for our results we scale up XMT processors one for one with full system nodes.

@@

\subsection{Applications}
\TODO{
    Traversal of an unbalanced, unpartitioned tree captures the foundations for implementing most forms of irregular parallelism. 
Recursion is found in every irregular divide and conquer algorithm.
Loops, even regular ones, must be dynamically scheduled to get good
load balance when run on a large, non-uniform system. Nested
parallelism involves fine-grain creation of dynamic amounts of work.
Dataflow..
}

\paragraph{Unbalanced tree search in-memory (UTS-Mem)}
Unbalanced Tree Search (UTS) is a benchmark
for evaluating the programmability and performance of systems for
parallel applications that require dynamic load balancing
\cite{Olivier:uts2006}. It involves traversing an unbalanced implicit
tree: at each vertex, its number of children is sampled from some
probability distribution, and this number of new nodes are added to a
work queue to be visited. While this benchmark captures irregular,
dynamic parallelism, we want to evaluate performance on algorithms
with irregular memory access patterns.

We augment UTS by using the existing traversal code to
create a large tree in memory, and then we traverse the in-memory
tree. We call this benchmark UTS-Mem, and the timed portion is this
traversal of the in-memory tree. This in-memory traversal is
unknowledgeable of the tree structure beforehand.
